Field-informed Reinforcement Learning for Learning Large-Scale Collective Tasks

### 2023-05-18 paper.tex

The \ac{GNN} versions demonstrate the capability to transfer the learned experience to the test phase whereas the MLP version fails to generalize. 
%
It is worth noting that, in the static configuration, 
 once the desired configuration is achieved in the static case, 
 the nodes cease to move, maintaining the found configuration.
%
Interesting observations arise when we use different scenarios from the training phase. 
 Starting from a division into two zones, 
 we notice that even in this case, the informed version finds a better configuration than the GNN counterpart. 
%
It exhibits both higher overall coverage and manages to divide the system into two equally covered parts. 
%
In contrast, the uninformed version also maintains a fair division but with significantly different coverage between the two parts, 
 indicating an imbalance among the nodes in one of the zones.
%
Finally, the moving scenario emphasizes how the informed version generates a more robust policy for new scenarios. 
 Indeed, we observe that the informed version maintains higher coverage and a greater number of agents on the target phenomenon compared to the other two solutions.
%
The uninformed GNN version, however, 
 fails to generalize its movement behaviour, as evidenced by the simulations where the agents, once reaching the target zone, stop moving due to tracking issues.

In conclusion, the results demonstrate how the proposed idea can generate more robust controllers. 
 By guiding information flow in \acp{GNN}, 
 we improve learning efficiency and alleviate the challenge of encoding relevant information. 
%
Nevertheless, we acknowledge the crucial role of \acp{GNN}. 
 Our modified version of DQN, combined with \acp{GNN}, 
 enables the discovery of robust behaviours in a few episodes, 
 which is challenging to capture with \acp{MLP} combined with \ac{DQN}, even if we use field information.
 \lukas{add short discussion on 'new' figure 6}
\begin{figure}
  \centering
  \hfill
  \begin{subfigure}[b]{0.96\linewidth}
    \includegraphics[width=\linewidth]{imgs/coverage-in-time}
    \caption{Average coverage for each episode in training }
    \label{fig:coverage}
  \end{subfigure}\\
  \hfill
  \begin{subfigure}[b]{0.99\linewidth}
    \includegraphics[width=\linewidth]{imgs/reward-in-time}
    \caption{Reward during training}
    \label{fig:reward}
  \end{subfigure}
  \caption{Training results of field-informed \ac{RL}. 
  Our solution is able to cover the phenomenon better than the baselines and it reaches a higher reward. 
  }
  \label{fig:training}
\end{figure}
\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/start.png}
      \caption{Start}
      \label{fig:initial}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/after.png}
      \caption{After 50 steps}
      \label{fig:after}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/end.png}
      \caption{After 200 steps}
      \label{fig:end}
  \end{subfigure}
  \caption{Qualitative test results of the proposed approach. 
  The agents go towards the phenomenon and try to cover it without collapsing into a single point.
  the grey area represents the area covered by the robots (last figure on the right).}
  \label{fig:simulations}
\end{figure}

>>>>

The \ac{GNN} versions demonstrate the capability to transfer the learned experience to the test phase whereas the MLP version fails to generalize. 
%
It is worth noting that, in the static configuration, 
 once the desired configuration is achieved in the static case, 
 the nodes cease to move, maintaining the found configuration.
%
Interesting observations arise when we use different scenarios from the training phase. 
 Starting from a division into two zones, 
 we notice that even in this case, the informed version finds a better configuration than the GNN counterpart. 
%
It exhibits both higher overall coverage and manages to divide the system into two equally covered parts.
%
Indeed, observing the \Cref{fig:resCoverage}, we notice that the informed version maintains a balanced coverage between the two zones, 
 with a difference of less than 5\% between the two parts, maintaining a fair division of the phenomenon.
%
In contrast, the uninformed version also maintains a fair division but with significantly different coverage between the two parts, 
 indicating a wrong placement among the nodes in one of the zones.
%
This is a consequence of the uninformed version's inability to encode the necessary information to divide the agents into two zones, therefore it is not able to generalise.
%
Finally, the moving scenario emphasizes how the informed version generates a more robust policy for new scenarios. 
 Indeed, we observe that the informed version maintains higher coverage and a greater number of agents on the target phenomenon compared to the other two solutions.
%
The uninformed GNN version, however, 
 fails again to generalize its movement behaviour, as evidenced by the simulations where the agents, once reaching the target zone, stop moving due to tracking issues.

In conclusion, the results demonstrate how the proposed idea can generate more robust controllers. 
 By guiding information flow in \acp{GNN}, 
 we improve learning efficiency and alleviate the challenge of encoding relevant information. 
%
Nevertheless, we acknowledge the crucial role of \acp{GNN}. 
 Our modified version of DQN, combined with \acp{GNN}, 
 enables the discovery of robust behaviours in a few episodes, 
 which is challenging to capture with \acp{MLP} combined with \ac{DQN}, even if we use field information.
 \lukas{add short discussion on 'new' figure 6}
\begin{figure}
  \centering
  \hfill
  \begin{subfigure}[b]{0.96\linewidth}
    \includegraphics[width=\linewidth]{imgs/coverage-in-time}
    \caption{Average coverage for each episode in training }
    \label{fig:coverage}
  \end{subfigure}\\
  \hfill
  \begin{subfigure}[b]{0.99\linewidth}
    \includegraphics[width=\linewidth]{imgs/reward-in-time}
    \caption{Reward during training}
    \label{fig:reward}
  \end{subfigure}
  \caption{Training results of field-informed \ac{RL}. 
  Our solution is able to cover the phenomenon better than the baselines and it reaches a higher reward. 
  }
  \label{fig:training}
\end{figure}
\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/start.png}
      \caption{Start}
      \label{fig:initial}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/after.png}
      \caption{After 50 steps}
      \label{fig:after}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\linewidth}
      \includegraphics[width=\textwidth]{imgs/end.png}
      \caption{After 200 steps}
      \label{fig:end}
  \end{subfigure}
  \caption{Qualitative test results of the proposed approach. 
  The agents go towards the phenomenon and try to cover it without collapsing into a single point.
  the grey area represents the area covered by the agents (last figure on the right).}
  \label{fig:simulations}
\end{figure}
